Coherent Hyper-Raman Four Wave Mixing Vibrational Spectroscopy

### 2024-07-20 paper.tex
2d figure fixed

Not all CMDS methods are used to dissect intramolecular couplings.\cite{Shen1987_CPL}
The most well known example of such a spectroscopy is vibrational sum frequency generation (vSFG), a three wave mixing technique used to probe vibrational structure and dynamics at interfaces. \cite{Shen1987_CPL, RN224}
Non-parametric, hyper difference frequency generation (hyper-DFG, or HDFG) spectroscopy, similar to difference frequency generation (\autoref{fig:comparisonwmel}), is a four wave mixing method which, when detuned from  electronic resonances, does not probe intramolecular couplings. 
HDFG was the first infrared CMDS method to successfully discriminate against non-resonant background.\cite{RN351, RN352}
HDFG has also been referred to as singly vibrationally enhanced (SIVE) spectroscopy. \cite{RN351}
In HDFG, an infrared pulse is resonant with a vibrational mode, and two other input pulses are used to stimulate a scattering process.
HDFG was documented long ago to have characteristics similar to both infrared spectroscopy and spontaneous hyper-Raman scattering. \cite{RN352}
Hyper-Raman scattering is the two photon analogue to Raman scattering. \cite{Cyvin1965, Terhune1965, Kozich2007}
Hyper-Raman scattering cross sections are commonly several orders of magnitude smaller than Raman scattering cross sections, making hyper-Raman a difficult process to detect.\cite{RN515, Kelley2010} 
The different selection rules of hyper-Raman scattering make it a unique alternative to Raman scattering to understand vibrational spectra and vibronic coupling in molecular systems. \cite{Olson2018}
Some years ago, Cho et al. proposed parametric six wave mixing techniques as the coherent analogue of hyper-Raman scattering. \cite{Cho1997, Cho1998}
However, since most six wave mixing methods cascade into four wave mixing processes and their homodyned output would be dependent upon the square magnitude of the hyper-Raman scattering cross section, \cite{RN243, RN515, Cho1997, Cho2000_Cascade} it is preferable to investigate four wave mixing analogues to hyper-Raman scattering.

@@

\subsection{Placzek Method}
We first investigate the selection rules through a Placzek type formalism.
Contracting over the virtual states forms the hyper-Raman hyperpolarizability $\beta$.\cite{Long1970} 
Assuming $\omega_2$, $\omega_3$ are significantly detuned from any resonance (\autoref{fig:comparisonwmel}),\cite{Placzek1934, Long1970} \autoref{sivegamma} is written as 
\begin{equation}\label{sivebeta}
	\gamma_{ijkl}^{vg} =	-\frac{1}{\varepsilon_0} \frac{1}{4D \hbar}\frac{\beta^{vg}_{ijk} \mu^{gv}_{l}}{\Delta_{gv}} \rho_{gg}
\end{equation}
It is important to note that the symmetry properties of $\beta_{ijk}$ change dependent upon whether $\omega_2$ and $\omega_3$ are degenerate. \cite{Denisov1986}
Taylor expanding the dipole and first hyperpolarizability operators in terms of an n$^{\text{th}}$ reduced normal mode coordinate $Q_n$ about equilibrium as\cite{Long1970, Shen90}
\begin{subequations}
	\begin{equation}
		\mu_l = \mu_{l,0} + \left(\frac{\partial \mu_l}{\partial Q_n}\right)_0 Q_n + \order{{Q_m Q_n}}
	\end{equation}
	\begin{equation}
		\beta_{ijk} = \beta_{ijk,0} + \left(\frac{\partial \beta_{ijk}}{\partial Q_n}\right)_0 Q_n + \order{{Q_m Q_n}}
	\end{equation}
\end{subequations}
and substituting into \autoref{sivebeta} gives the HDFG hyperpolarizability to $\order{Q_n}$ as \begin{equation}\label{SIVEselection}
	\gamma_{ijkl}^{vg} =	-\frac{1}{\varepsilon_0} \frac{1}{8D m_n \omega_{vg}}  \frac{1}{{\Delta_{gv}}} \ \left(\frac{\partial \beta^{vg}_{ijk}}{\partial Q_n}\right)_0 \left({\frac{\partial \mu^{gv}_{l}}{\partial Q_n}}\right)_0  \rho_{gg}
\end{equation}
where $\mel{v}{Q_n}{g} = \sqrt{\frac{\hbar}{2\omega_{vg}}}$.  \cite{RN230}
Since this expression is non-zero in the harmonic oscillator limit, HDFG output is allowed for harmonic transitions. 
Similar to SFG, HDFG does not demand anharmonic ground state potential energy surfaces for spectral output. \cite{Shen94, Cho2000}
This selection rule is generally valid for any HDFG or hyper sum frequency generation (HSFG) process in the $-\vec{k}_1 + \vec{k}_2  + \vec{k}_3$ and $\vec{k}_1 + \vec{k}_2  + \vec{k}_3$ phasematching geometries when $\omega_2$ and $\omega_3$ are sufficiently detuned from any resonance.

@@

The impact of vibronic coupling in the 2D spectrum is made clear by using a simple model.
Using a two-well system described by $H_g = p^2/2m + 1/2 \hbar \omega Q^2$ and $H_e = p^2/2m + 1/2 \hbar \omega (Q-\Delta)^2 +\hbar \omega_{eg}$, the A and B contributions can be easily evaluated using Franck-Condon and Herzberg-Teller integrals tabulated elsewhere. \cite{Carlson1988thesis} 
A simulated spectrum highlighting the contributions of these terms (\autoref{fig:doubres_spec}) on potential wells with $\Delta = 0.5$ using values listed in \autoref{T:hdfgvalues} dissects the $A$ and $B$ contributions to the DR-SIVE spectrum.
%stimulated hyper Raman spectrum -- relationship
\begin{figure}[!htbp]
	\centering
	\includegraphics[width=3.375in]{drsive_spectrum.png}
	\caption{DR-HDFG $\gamma_{ijkl}$ spectrum for a simple two harmonic well system.
		(a) Potential energy surface, (b) HDFG spectrum for $\Delta = 0.5$.}
	\label{fig:doubres_spec}
\end{figure}

>>>>

Not all CMDS methods are used to dissect intramolecular couplings.\cite{Shen1987_CPL}
The most well known example of such a spectroscopy is vibrational sum frequency generation (vSFG), a three wave mixing technique used to probe vibrational structure and dynamics at interfaces. \cite{Shen1987_CPL, RN224}
Non-parametric, hyper difference frequency generation (hyper-DFG, or HDFG) spectroscopy, similar to difference frequency generation (\autoref{fig:comparisonwmel}), is a four wave mixing method which, when detuned from  electronic resonances, does not probe intramolecular couplings. 
HDFG was the first infrared CMDS method to successfully discriminate against non-resonant background.\cite{RN351, RN352}
HDFG has also been referred to as singly vibrationally enhanced (SIVE) spectroscopy. \cite{RN351}
In HDFG, an infrared pulse is resonant with a vibrational mode, and two other input pulses are used to stimulate a scattering process.
HDFG was documented long ago to have characteristics similar to both infrared spectroscopy and spontaneous hyper-Raman scattering. \cite{RN352}
Hyper-Raman scattering is the two photon analogue to Raman scattering. \cite{Cyvin1965, Terhune1965, Andrews1978}
Hyper-Raman scattering cross sections are commonly several orders of magnitude smaller than Raman scattering cross sections, making hyper-Raman a difficult process to detect.\cite{RN515, Kelley2010} 
However, unlike Raman spectroscopy, all infrared active modes are hyper-Raman active, which allows infrared active modes to be probed through an inelastic scattering event. \cite{Andrews1978}
The different selection rules of hyper-Raman scattering make it a unique alternative to Raman scattering to understand vibrational spectra and vibronic coupling in molecular systems. \cite{Olson2018}
Some years ago, Cho et al. proposed parametric six wave mixing techniques as the coherent analogue of hyper-Raman scattering. \cite{Cho1997, Cho1998}
However, since most six wave mixing methods cascade into four wave mixing processes and their homodyned output would be dependent upon the hyper-Raman scattering cross section, \cite{RN243, RN515, Cho1997, Cho2000_Cascade} it is preferable to investigate four wave mixing analogues to hyper-Raman scattering.

@@

\subsection{Placzek Method}
We first investigate the selection rules through a Placzek type formalism.
Contracting over the virtual states forms the hyper-Raman hyperpolarizability $\beta$.\cite{Long1970} 
Assuming $\omega_2$, $\omega_3$ are significantly detuned from any resonance (\autoref{fig:comparisonwmel}),\cite{Placzek1934, Long1970, Altmann1982} \autoref{sivegamma} is written as 
\begin{equation}\label{sivebeta}
	\gamma_{ijkl}^{vg} =	-\frac{1}{\varepsilon_0} \frac{1}{4D \hbar}\frac{\beta^{vg}_{ijk} \mu^{gv}_{l}}{\Delta_{gv}} \rho_{gg}
\end{equation}
All infrared active transitions are hyper-Raman active, making HDFG allowed for any infrared active transition. \cite{Andrews1978}
It is important to note that the symmetry properties of $\beta_{ijk}$ change dependent upon whether $\omega_2$ and $\omega_3$ are degenerate. \cite{Denisov1986, Kozich2007}
Taylor expanding the dipole and first hyperpolarizability operators in terms of an n$^{\text{th}}$ reduced normal mode coordinate $Q_n$ about equilibrium as\cite{Long1970, Shen90}
\begin{subequations}
	\begin{equation}
		\mu_l = \mu_{l,0} + \left(\frac{\partial \mu_l}{\partial Q_n}\right)_0 Q_n + \order{{Q_m Q_n}}
	\end{equation}
	\begin{equation}
		\beta_{ijk} = \beta_{ijk,0} + \left(\frac{\partial \beta_{ijk}}{\partial Q_n}\right)_0 Q_n + \order{{Q_m Q_n}}
	\end{equation}
\end{subequations}
and substituting into \autoref{sivebeta} gives the HDFG hyperpolarizability to $\order{Q_n}$ as \begin{equation}\label{SIVEselection}
	\gamma_{ijkl}^{vg} =	-\frac{1}{\varepsilon_0} \frac{1}{8D m_n \omega_{vg}}  \frac{1}{{\Delta_{gv}}} \ \left(\frac{\partial \beta^{vg}_{ijk}}{\partial Q_n}\right)_0 \left({\frac{\partial \mu^{gv}_{l}}{\partial Q_n}}\right)_0  \rho_{gg}
\end{equation}
where $\mel{v}{Q_n}{g} = \sqrt{\frac{\hbar}{2\omega_{vg}}}$.  \cite{RN230}
Since this expression is non-zero in the harmonic oscillator limit, HDFG output is allowed for harmonic transitions. 
Similar to SFG, HDFG does not demand anharmonic ground state potential energy surfaces for spectral output. \cite{Shen94, Cho2000}
This selection rule is generally valid for any HDFG or hyper sum frequency generation (HSFG) process in the $-\vec{k}_1 + \vec{k}_2  + \vec{k}_3$ and $\vec{k}_1 + \vec{k}_2  + \vec{k}_3$ phasematching geometries when $\omega_2$ and $\omega_3$ are sufficiently detuned from any resonance.

@@

The impact of vibronic coupling in the 2D spectrum is made clear by using a simple model.
Using a two-well system described by $H_g = p^2/2m + 1/2 \hbar \omega Q^2$ and $H_e = p^2/2m + 1/2 \hbar \omega (Q-\Delta)^2 +\hbar \omega_{eg}$, the A and B contributions can be easily evaluated using Franck-Condon and Herzberg-Teller integrals tabulated elsewhere. \cite{Carlson1988thesis} 
A simulated spectrum highlighting the contributions of these terms (\autoref{fig:doubres_spec}) on potential wells with $\Delta = 0.5$ using values listed in \autoref{T:hdfgvalues} dissects the $A$ and $B$ contributions to the DR-SIVE spectrum.
%stimulated hyper Raman spectrum -- relationship; make clear that this spectrum highlights anything that ends in |gv> from the IR transition!
\begin{figure*}[!htbp]
	\centering
	\includegraphics[width=6.66in]{drsive_spectrum.png}
	\caption{DR-HDFG $\gamma_{ijkl}$ spectrum for a simple two harmonic well system.
		(a) Potential energy surface, (b) HDFG spectrum for $\Delta = 0.5$.}
	\label{fig:doubres_spec}
\end{figure*}
